---
title: 'PFAS and environmental justice interactive map'
subtitle: "beta version"
author: "Jahred Liddie, Dr. Laurel Schaider, Dr. Elsie Sunderland"
output:
  html_document:
    theme: journal
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
# load data and set up libraries
library(leaflet)
library(sf)
library(maps)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(htmlwidgets)

# PFAS data
CWS <- read.csv("../../Merged PFAS data/final_dat_FN.csv")
CWS.counties <- st_read("../../Merged PFAS data/CWS_formapping.geojson")
centroids <- st_read("../../Merged PFAS data/CWS_centroid.geojson")
ucmr.counties <- st_read("../../Merged PFAS data/UCMR_formapping.geojson")

CWS.counties <- st_transform(CWS.counties, crs = st_crs(centroids))
ucmr.counties <- st_transform(ucmr.counties, crs = st_crs(centroids))

# PFAS source data
  # MFTA
  military <- geojsonio::geojson_read("../../PFAS point source data/Military_2020MARCH04.geojson", what = "sp")
  military <- st_as_sf(military, )

  # airports
  airports <- readxl::read_excel("../../PFAS point source data/Part 139_cert_airports.xlsx")
  airports <- as.data.frame(airports)
  
  # epa stewardship sites
  epa <- readxl::read_excel("../../PFAS point source data/EPA 2010.2015 PFOA Stewardship Program sites.xlsx")
  
  # CWNS wastewater treatment plants
  WWTP <- read.csv("../../PFAS point source data/WWTP facility_details.csv")
  
  # sources for secondary analyses: landfills from LMOP
  landfills <- read.csv("../../PFAS point source data/Secondary analysis/landfills_ewg.csv")
  
  landfills.lmop <- read.csv("../../PFAS point source data/Secondary analysis/landfilllmopdata_clean.csv")

  CWS.states <- c("california", "colorado", "kentucky", "massachusetts", "michigan",
              "maryland", "new hampshire", "new jersey", "pennsylvania", "ohio",
              "vermont")
  
  ucmr.states <- c(tolower(ucmr.counties$state_name))
    
  # note: need to intersect military + airports with US boundary before mapping
  # also explore not showing points at first
  
  # county boundaries
  # CWS.counties <- map_data("county", regions = CWS.states)
  # ucmr.counties <- map_data("county", regions = ucmr.states)
  #   
  # CWS.counties <- st_as_sf(CWS.counties, coords = c("long", "lat"),
  #                          crs = st_crs(centroids))
  # ucmr.counties <- st_as_sf(ucmr.counties, coords = c("long", "lat"),
  #                           crs = st_crs(centroids))

knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning = FALSE, results='hide', echo = FALSE}
# clean up source data
  # ignoring WWTPs for now because there aren't points available
  # # add leading zero to HUC codes if they are under 8 characters
  # WWTP$Watershed.HUC <- ifelse(str_length(WWTP$Watershed.HUC) < 8, 
  #                               paste("0", WWTP$Watershed.HUC, sep=""),
  #                               WWTP$Watershed.HUC)
  # 
  # WWTP.HUCs <- WWTP %>% group_by(Watershed.HUC, Watershed.Name) %>% 
  #   dplyr::summarise(WWTP_count = n(), 
  #          WWTP_existingtotalflow_Mgal.d = sum(as.numeric(Existing.Total.Flow..Mgal.d.))) %>% 
  #   ungroup()
  # 
  # WWTP.HUCs <- WWTP.HUCs %>% dplyr::rename(
  #   HUC_CODE = Watershed.HUC,
  #   HUC_NAME = Watershed.Name
  # ) 

  ### Military sites ###
  # note: 9 of those dropped from here are in AK
  # st_crs(military) == st_crs(centroids) # check if true
  # military$CWS.ind <- st_intersects(military, counties)
  
  ### EPA Stewardship sites ###
  # separate into lat and lon
  epa <- separate(epa, col = "Coordinates", into = c("lat", "lon"), sep = ",")
  
  # convert to sf object
  epa <- st_as_sf(epa, coords = c("lon", "lat"), crs = "NAD83")
  epa <- st_transform(epa, crs = st_crs(centroids))
  epa$CWS.ind <- epa$State_ABV %in% unique(CWS.counties$state)
  epa$ucmr.ind <- epa$State_ABV %in% unique(ucmr.counties$state)
  
  ### Airports ###
  # clean up data
  airports <- airports %>% dplyr::select(LocationID, Region, State, StateName, County, 
                                              CountyState, City, FacilityName, ARPLatitude, 
                                              ARPLatitudeS, ARPLongitude, ARPLongitudeS)
  
  airports <- separate(airports, col = "ARPLatitude", 
                       into = c("lat.deg", "lat.min", "lat.sec.dir"), sep = "-")
  
  airports <- separate(airports, col = "ARPLongitude", 
                       into = c("lon.deg", "lon.min", "lon.sec.dir"), sep = "-")
  
  # create columns without direction
  airports <- airports %>% 
    mutate(lat.sec = as.numeric(substr(lat.sec.dir, 1, nchar(lat.sec.dir)-1)),
           lon.sec = as.numeric(substr(lon.sec.dir, 1, nchar(lon.sec.dir)-1))) 
  
  # calculate lat and lon
  airports <- airports %>% 
    mutate(
      lat = ifelse(grepl("N", lat.sec.dir), 
                        as.numeric(lat.deg) + as.numeric(lat.min)/60 + lat.sec/3600,
                        -1 * (as.numeric(lat.deg) + as.numeric(lat.min)/60 + lat.sec/3600)),
      lon = ifelse(grepl("E", lon.sec.dir), 
                        as.numeric(lon.deg) + as.numeric(lon.min)/60 + lon.sec/3600,
                        -1 * (as.numeric(lon.deg) + as.numeric(lon.min)/60 + lon.sec/3600)),
    )
  
  # convert to sf object
  airports <- st_as_sf(airports, 
                            coords = c("lon", "lat"), crs = st_crs(centroids))
  airports$CWS.ind <- airports$State %in% unique(CWS.counties$state)
  airports$ucmr.ind <- airports$State %in% unique(ucmr.counties$state)
  
  # remove airports outside of 48 states + HI, AK
  outside50 <- c("'PPG", "'Z08", "'FAQ", "'MDY", "'STX", "'STT", "'PSE", 
                 "'SJU", "'BQN")
  
  airports <- airports %>% filter(!(LocationID %in% outside50))
  
    # now with LMOP data
  landfills.lmop <- landfills.lmop %>% 
    select(Landfill.ID, Landfill.Name, 
           lon = Longitude,
           lat = Latitude) %>% 
    filter(!is.na(lat) & !is.na(lon)) %>%
    # drop missing coords
    st_as_sf(coords = c("lon", "lat"), crs = st_crs(centroids))
  
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# clean up centroids; take average of systems with multiple locations
centroids.nosf <- centroids %>%
  st_coordinates() %>%
   cbind(st_drop_geometry(centroids),.)

centroids.nosf <- centroids.nosf %>%
  group_by(PWSID) %>%
  mutate(X = mean(X),
         Y = mean(Y))

centroids.nosf <- unique(centroids.nosf)

centroids.nosf <- st_as_sf(centroids.nosf, coords = c("X", "Y"), 
                           crs = st_crs(centroids))

# overwrite now
centroids <- centroids.nosf

centroids <- left_join(centroids, CWS)

```


```{r, echo=FALSE, include=FALSE}
# demographic centering (by state)
CWS.counties <- CWS.counties %>%
  group_by(state) %>%
  mutate(mean.Hisp = mean(percHisp),
         mean.Black = mean(percBlack),
         mean.White = mean(percWhite),
         mean.poverty = mean(poverty)
         ) %>% ungroup()

CWS.counties <- CWS.counties %>%
  mutate(center.Hisp = percHisp - mean.Hisp,
         center.Black = percBlack - mean.Black,
         center.White = percWhite - mean.White,
         center.poverty = poverty - mean.poverty
         ) %>% ungroup()

ucmr.counties <- ucmr.counties %>%
  group_by(state) %>%
  mutate(mean.Hisp = mean(percHisp),
         mean.Black = mean(percBlack),
         mean.White = mean(percWhite),
         mean.poverty = mean(poverty)
         ) %>% ungroup()

ucmr.counties <- ucmr.counties %>%
  mutate(center.Hisp = percHisp - mean.Hisp,
         center.Black = percBlack - mean.Black,
         center.White = percWhite - mean.White,
         center.poverty = poverty - mean.poverty
         ) %>% ungroup()

ucmr.counties$urban <- ifelse(ucmr.counties$percurban >= 0.5, "Urban", "Rural")
CWS.counties$urban <- ifelse(CWS.counties$percurban >= 0.5, "Urban", "Rural")

```

```{r, echo=FALSE, warning=FALSE}
# function for making palettes
makepalette.f <- function(variable){

  variable_palette <- colorNumeric(palette = "Blues", CWS.counties$variable)
  
  variable_palette(variable)
}

# function for popups
popup.f <- function(dataset) {
  
  paste("<B>Demographics</B>", "<br/>",
        "% Hispanic: ", 
        round(dataset$percHisp, 0), "%", "<br/>",
        "% Black non-Hispanic: ",
        round(dataset$percBlack, 0), "%", "<br/>",
        "% White non-Hispanic: ",
        round(dataset$percWhite, 0), "%", "<br/>",
        "% living under the federal poverty line: ",
        round(dataset$poverty, 0), "%", "<br/>",
        "Urban or rural (≥50% residents in urban areas): ", dataset$urban, "<br/>",
        "<br/>",
        "<B>PFAS data (% of systems with detectable levels) </B>", "<br/>",
        "PFOA: ", 
        round(dataset$PFOA.avg, 0), "%", "<br/>",
        "PFOS: ", 
        round(dataset$PFOS.avg, 0), "%", "<br/>",
        "PFNA: ", 
        round(dataset$PFNA.avg, 0), "%", "<br/>",
        "PFHxS: ", 
        round(dataset$PFHxS.avg, 0), "%", "<br/>",
        "PFBS: ", 
        round(dataset$PFBS.avg, 0), "%", "<br/>",
        "At least 1 PFAS: ", 
        round(dataset$Any.avg, 0), "%", "<br/>",
        "Number of community water systems in dataset: ",
        dataset$num.CWS,
        sep = "")
}

  CWS.counties$urban <- as.factor(CWS.counties$urban)
  pal.urban <- colorFactor(palette = c("blue", "darkgreen"), CWS.counties$urban)
  
  # color ramps for centered variables
  rc1 <- colorRampPalette(colors = c("red", "white"), space = "Lab")(round(abs(min(CWS.counties$center.Hisp)), 0))
  
  ## Make vector of colors for values larger than 0
  rc2 <- colorRampPalette(colors = c("white", "blue"), space = "Lab")(round(abs(max(CWS.counties$center.Hisp)), 0))
  
  rampcols <- c(rc1, rc2)
  
  pal.center.Hisp <- colorNumeric(palette = rampcols, domain = CWS.counties$center.Hisp)
  
  # color ramps for centered variables
  rc1 <- colorRampPalette(colors = c("red", "white"), space = "Lab")(round(abs(min(CWS.counties$center.Black)), 0))
  
  ## Make vector of colors for values larger than 0
  rc2 <- colorRampPalette(colors = c("white", "blue"), space = "Lab")(round(abs(max(CWS.counties$center.Black)), 0))
  
  rampcols <- c(rc1, rc2)
  
  pal.center.Black <- colorNumeric(palette = rampcols, domain = CWS.counties$center.Black)
  
    # color ramps for centered variables
  rc1 <- colorRampPalette(colors = c("red", "white"), space = "Lab")(round(abs(min(CWS.counties$center.poverty)), 0))
  
  ## Make vector of colors for values larger than 0
  rc2 <- colorRampPalette(colors = c("white", "blue"), space = "Lab")(round(abs(max(CWS.counties$center.poverty)), 0))
  
  rampcols <- c(rc1, rc2)
  
  pal.center.poverty <- colorNumeric(palette = rampcols, domain = CWS.counties$center.poverty)

  
  # individual systems processing for map
centroids_map <- subset(centroids, !is.na(Any_detect))
centroids_map$Any_detect_map <- ifelse(centroids_map$Any_detect == 1,
                                       "≥1 PFAS detected (>5 ng/L)", "No PFAS detected")
  pal.PFAS <- colorFactor(palette = c('darkred', 'darkgreen'), 
                          domain = centroids_map$Any_detect_map)
  
```

# Statewide sampling dataset

The interactive map below shows PFAS drinking water data, sociodemographics, and datasets on confirmed and likely PFAS contamination sources obtained as part of **"Sociodemographic Determinants of PFAS Contamination in United States Community Water Supplies"**. These data were collated from statewide sampling campaigns from community water systems in 11 states over 2016-2021.  

Percent Black and White refer to non-Hispanic Black and White residents, respectively. To analyze within-state demographic trends, polygon layers using "centered" demographics (substracting the state-specific mean) are also provided. PFAS data for individual water systems can also be viewed - these should be viewed with caution, as water systems may serve multiple / non-adjacent areas and geocoded locations may be inexact. Concentrations for the statewide sampling data refer to measurement of that PFAS at least once over 5 nanograms per liter. Additional information on all data are available in the associated paper. 

To change layers, click on the panel in the bottom right.

```{r, echo=FALSE, warning=FALSE}

limits <- st_bbox(ucmr.counties)

map1 <- leaflet(CWS.counties) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB.Positron") %>%
  
      # PFAS sources
  addCircleMarkers(data = epa, radius = 3, group = "EPA Stewardship sites",
                   weight = 3, color = "black", 
                   fillColor = "goldenrod", fillOpacity = 1) %>%
  addCircleMarkers(data = military, radius = 3, group = "Military fire-training areas",
                 weight = 3, color = "black", fillColor = "purple", fillOpacity = 1) %>%
  addCircleMarkers(data = airports, radius = 3, group = "Airports (AFFF certified)",
                 weight = 3, clusterOptions = markerClusterOptions(),
                 color = "black", fillColor = "green", fillOpacity = 1) %>%
  
  # CWS centroids
  addCircleMarkers(data = centroids_map, radius = 3, group = "Individual community water systems",
                   weight = 3, clusterOptions = markerClusterOptions(),
                   fillOpacity = 1, color = ~pal.PFAS(Any_detect_map),
                   popup = centroids_map$Any_detect_map) %>%
  
  # other providers: https://leaflet-extras.github.io/leaflet-providers/preview/
  # PFAS variables
  addPolygons(fillColor = ~makepalette.f(PFOA.avg), stroke = TRUE, 
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "Detection of PFOA",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties))  %>%
  addPolygons(fillColor = ~makepalette.f(PFOS.avg), stroke = TRUE, 
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "Detection of PFOS",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%
  addPolygons(fillColor = ~makepalette.f(Any.avg), stroke = TRUE, 
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "Detection of ≥1 PFAS",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%

  addPolygons(fillColor = ~makepalette.f(percHisp), stroke = TRUE, 
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% Hispanic residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%

  addPolygons(fillColor = ~makepalette.f(percBlack),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% Black residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%

  addPolygons(fillColor = ~makepalette.f(percWhite),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% White residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%
  
  addPolygons(fillColor = ~pal.center.Hisp(center.Hisp), stroke = TRUE,
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% Hispanic residents (centered)",
              highlightOptions = highlightOptions(fillColor = "darkblue",
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%
  addPolygons(fillColor = ~pal.center.Black(center.Black), stroke = TRUE,
            color = "black", opacity = 1, weight = 1,
            fillOpacity = 1, group = "% Black residents (centered)",
            highlightOptions = highlightOptions(fillColor = "darkblue",
                                                fillOpacity = 0.9),
            label = CWS.counties$NAME,
            popup = popup.f(dataset = CWS.counties)) %>%
  addPolygons(fillColor = ~pal.center.poverty(center.poverty), stroke = TRUE,
            color = "black", opacity = 1, weight = 1,
            fillOpacity = 1, group = "% residents living under federal poverty line (centered)",
            highlightOptions = highlightOptions(fillColor = "darkblue",
                                                fillOpacity = 0.9),
            label = CWS.counties$NAME,
            popup = popup.f(dataset = CWS.counties)) %>%
  addPolygons(fillColor = ~makepalette.f(poverty),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group =  "% residents living under federal poverty line",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup.f(dataset = CWS.counties)) %>%
  addPolygons(fillColor = ~pal.urban(urban),
            stroke = TRUE, color = "black", opacity = 1, weight = 1,
            fillOpacity = 1, group =  "Urban or rural area",
            highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                fillOpacity = 0.9),
            label = CWS.counties$NAME,
            popup = popup.f(dataset = CWS.counties)) %>%

  # addCircleMarkers(data = landfills.lmop, radius = 3, group = "Municipal landfills",
  #                weight = 3,
  #                color = "black", fillColor = "black", fillOpacity = 1) %>%
  addLayersControl(
    position = "bottomright",
    baseGroups  = c("% Hispanic residents", "% Hispanic residents (centered)",
                    "% Black residents", "% Black residents (centered)", "% White residents",
                    "% residents living under federal poverty line",
                    "% residents living under federal poverty line (centered)",
                    "Urban or rural area",
                    "Detection of PFOA", "Detection of PFOS", "Detection of ≥1 PFAS"),
    overlayGroups = c("EPA Stewardship sites", "Military fire-training areas",
                      "Airports (AFFF certified)",
                      "Individual community water systems"),
    options = layersControlOptions(collapsed = TRUE)) %>%
    setMaxBounds(lng1 = as.numeric(limits[1]) - 20,
                 lat1 = as.numeric(limits[2]) - 20,
                 lng2 = as.numeric(limits[3]) + 20,
                 lat2 = as.numeric(limits[4]) + 20
               )

setView(map1 %>% hideGroup(c("EPA Stewardship sites", "Military fire-training areas",
                     "Airports (AFFF certified)", "Individual community water systems")),
        lat = 40, lng = 42, zoom = 3)


```


```{r, echo=FALSE, warning=FALSE, eval=FALSE}

popup_short.f <- function(dataset) {
  
  paste("<B>Demographics</B>", "<br/>",
        "% Hispanic: ", 
        round(dataset$percHisp, 0), "%", "<br/>",
        "% Black non-Hispanic: ",
        round(dataset$percBlack, 0), "%", "<br/>",
        "% White: ",
        round(dataset$percWhite, 0), "%", "<br/>",
        "% living under the federal poverty line: ",
        round(dataset$poverty, 0), "%", "<br/>", "<br/>",
        sep = "")
}

map2 <- leaflet(CWS.counties) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB.Positron") %>%
  
  # other providers: https://leaflet-extras.github.io/leaflet-providers/preview/
  # PFAS variables
  # addPolygons(fillColor = ~makepalette.f(PFOA.avg), stroke = TRUE, 
  #             color = "black", opacity = 1, weight = 1,
  #             fillOpacity = 1, group = "Detection of PFOA",
  #             highlightOptions = highlightOptions(fillColor = "darkblue", 
  #                                                 fillOpacity = 0.9),
  #             label = CWS.counties$NAME,
  #             popup = popup.f(dataset = CWS.counties))  %>%
  # addPolygons(fillColor = ~makepalette.f(PFOS.avg), stroke = TRUE, 
  #           color = "black", opacity = 1, weight = 1,
  #           fillOpacity = 1, group = "Detection of PFOS",
  #           highlightOptions = highlightOptions(fillColor = "darkblue", 
  #                                               fillOpacity = 0.9),
  #           label = CWS.counties$NAME,
  #           popup = popup.f(dataset = CWS.counties)) %>%
  # addPolygons(fillColor = ~makepalette.f(Any.avg), stroke = TRUE, 
  #         color = "black", opacity = 1, weight = 1,
  #         fillOpacity = 1, group = "Detection of ≥1 PFAS",
  #         highlightOptions = highlightOptions(fillColor = "darkblue", 
  #                                             fillOpacity = 0.9),
  #         label = CWS.counties$NAME,
  #         popup = popup.f(dataset = CWS.counties)) %>%

    addPolygons(fillColor = ~makepalette.f(percHisp), stroke = TRUE, 
                color = "black", opacity = 1, weight = 1,
                fillOpacity = 1, group = "% Hispanic residents",
                highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                    fillOpacity = 0.9),
                label = CWS.counties$NAME,
                popup = popup_short.f(dataset = CWS.counties)) %>%
  
    addPolygons(fillColor = ~makepalette.f(percBlack),
                stroke = TRUE, color = "black", opacity = 1, weight = 1,
                fillOpacity = 1, group = "% Black residents",
                highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                    fillOpacity = 0.9),
                label = CWS.counties$NAME,
                popup = popup_short.f(dataset = CWS.counties)) %>%
  
    addPolygons(fillColor = ~makepalette.f(percWhite),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% White residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup_short.f(dataset = CWS.counties)) %>%
    addPolygons(fillColor = ~makepalette.f(center.Hisp), stroke = TRUE,
        color = "black", opacity = 1, weight = 1,
        fillOpacity = 1, group = "% Hispanic residents (centered)",
        highlightOptions = highlightOptions(fillColor = "darkblue",
                                            fillOpacity = 0.9),
        label = CWS.counties$NAME,
        popup = popup_short.f(dataset = CWS.counties)) %>%
    addPolygons(fillColor = ~makepalette.f(poverty),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group =  "% residents living under federal poverty line",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = CWS.counties$NAME,
              popup = popup_short.f(dataset = CWS.counties)) %>%
  
    # PFAS water data
  addCircleMarkers(data = centroids, radius = 2, group = "CWS",
                   weight = 2, color = ~ ifelse(PFOA_detect == 1, "red", 
                                                ifelse(PFOA_detect == 0, "blue", "white")),
                   # clusterOptions = markerClusterOptions(maxClusterRadius = 200),
                   popup = ~ifelse(PFOA_detect==1, "PFOA detected > 5 ng/L", 
                                   ifelse(PFOA_detect == 0, "PFOA not detected < 5 ng/L", "N/A")),
                   fillOpacity = 1) %>%
  
    # PFAS sources
  addCircleMarkers(data = epa, radius = 3, group = "EPA Stewardship sites",
                   weight = 3, color = "black", 
                   fillColor = "goldenrod", fillOpacity = 1) %>%
  addCircleMarkers(data = military, radius = 3, group = "Military fire-training areas",
                 weight = 3, color = "black", fillColor = "purple", fillOpacity = 1) %>%
  addCircleMarkers(data = airports, radius = 3, group = "Airports (AFFF certified)",
                 weight = 3, clusterOptions = markerClusterOptions(),
                 color = "black", fillColor = "green", fillOpacity = 1) %>%
  
  addLayersControl(
    position = "bottomright",
    baseGroups  = c("% Hispanic residents", "% Hispanic residents (centered)",
                    "% Black residents", "% White residents",
                    "% residents living under federal poverty line"),
    overlayGroups = c("EPA Stewardship sites", "Military fire-training areas",
                      "Airports (AFFF certified)"),
    options = layersControlOptions(collapsed = TRUE)) %>%
    setMaxBounds(lng1 = as.numeric(limits[1]),
               lat1 = as.numeric(limits[2]) - 10,
               lng2 = as.numeric(limits[3]) + 10,
               lat2 = as.numeric(limits[4]) + 10
               )

setView(map2 %>% hideGroup(c("EPA Stewardship sites", 
                             "Military fire-training areas",
                             "Airports (AFFF certified)")),
        lat = 40, lng = 42, zoom = 3)

```


<!-- ## UCMR 3   -->

<!-- Data from the US Environmental Protection Agency's Unregulated Contaminant Monitoring Rule 3 (2013-2015) are shown in the interactive map below. Note that detection limits in this nationwide monitoring campaign were higher than in the more recent state sampling campaigns, ranging from 10-90 ng/L. -->

```{r,  echo=FALSE, warning=FALSE, eval=FALSE}

map3 <- leaflet(ucmr.counties) %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB.Positron") %>%
  
  # other providers: https://leaflet-extras.github.io/leaflet-providers/preview/
  # PFAS variables
  addPolygons(fillColor = ~makepalette.f(PFOA.avg), stroke = TRUE, 
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "Detection of PFOA (> 20 ng/L)",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = ucmr.counties$NAME,
              popup = popup.f(dataset = ucmr.counties))  %>%
  addPolygons(fillColor = ~makepalette.f(PFOS.avg), stroke = TRUE, 
            color = "black", opacity = 1, weight = 1,
            fillOpacity = 1, group = "Detection of PFOS (> 20 ng/L)",
            highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                fillOpacity = 0.9),
            label = ucmr.counties$NAME,
            popup = popup.f(dataset = ucmr.counties)) %>%
  addPolygons(fillColor = ~makepalette.f(Any.avg), stroke = TRUE, 
          color = "black", opacity = 1, weight = 1,
          fillOpacity = 1, group = "Detection of ≥1 PFAS",
          highlightOptions = highlightOptions(fillColor = "darkblue", 
                                              fillOpacity = 0.9),
          label = ucmr.counties$NAME,
          popup = popup.f(dataset = ucmr.counties)) %>%

  addPolygons(fillColor = ~makepalette.f(percHisp), stroke = TRUE, 
              color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% Hispanic residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = ucmr.counties$NAME,
              popup = popup.f(dataset = ucmr.counties)) %>%

  addPolygons(fillColor = ~makepalette.f(percBlack),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% Black residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = ucmr.counties$NAME,
              popup = popup.f(dataset = ucmr.counties)) %>%
  
    addPolygons(fillColor = ~makepalette.f(percWhite),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group = "% White residents",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = ucmr.counties$NAME,
              popup = popup.f(dataset = ucmr.counties)) %>%
    addPolygons(fillColor = ~makepalette.f(center.Hisp), stroke = TRUE,
        color = "black", opacity = 1, weight = 1,
        fillOpacity = 1, group = "% Hispanic residents (centered)",
        highlightOptions = highlightOptions(fillColor = "darkblue",
                                            fillOpacity = 0.9),
        label = ucmr.counties$NAME,
        popup = popup.f(dataset = ucmr.counties)) %>%
    addPolygons(fillColor = ~makepalette.f(poverty),
              stroke = TRUE, color = "black", opacity = 1, weight = 1,
              fillOpacity = 1, group =  "% residents living under federal poverty line",
              highlightOptions = highlightOptions(fillColor = "darkblue", 
                                                  fillOpacity = 0.9),
              label = ucmr.counties$NAME,
              popup = popup.f(dataset = ucmr.counties)) %>%
  
    # PFAS sources
  # addCircleMarkers(data = epa, radius = 3, group = "EPA Stewardship sites",
  #                  weight = 3, color = "black", 
  #                  fillColor = "goldenrod", fillOpacity = 1) %>%
  # addCircleMarkers(data = military, radius = 3, group = "Military fire-training areas",
  #                weight = 3, color = "black", fillColor = "purple", fillOpacity = 1) %>%
  # addCircleMarkers(data = airports, radius = 3, group = "Airports (AFFF certified)",
  #                weight = 3, clusterOptions = markerClusterOptions(),
  #                color = "black", fillColor = "green", fillOpacity = 1) %>%

  # addCircleMarkers(data = landfills.lmop, radius = 3, group = "Municipal landfills",
  #                weight = 3,
  #                color = "black", fillColor = "black", fillOpacity = 1) %>%
  addLayersControl(
    position = "bottomright",
    baseGroups  = c("% Hispanic residents", "% Hispanic residents (centered)",
                    "% Black residents", "% White residents",
                    "% residents living under federal poverty line",
                    "Detection of PFOA", "Detection of PFOS", "Detection of ≥1 PFAS"),
    overlayGroups = c("EPA Stewardship sites", "Military fire-training areas",
                      "Airports (AFFF certified)"),
    options = layersControlOptions(collapsed = TRUE)) %>%
    setMaxBounds(lng1 = as.numeric(limits[1]),
               lat1 = as.numeric(limits[2]) - 10,
               lng2 = as.numeric(limits[3]) + 10,
               lat2 = as.numeric(limits[4]) + 10
               )

setView(map3 %>% hideGroup(c("EPA Stewardship sites",
                             "Military fire-training areas",
                             "Airports (AFFF certified)")),
        lat = 40, lng = 42, zoom = 3)
```

